Causal Compatibility Inequalities Admitting of Quantum Violations in the Triangle Scenario

### 2016-11-11 documents/BrazilSlides/slides.tex
Marginal Problem foramlism

% \usepackage{pgfpages} % For viewing slides on the right
\usepackage[warningthreshold=0.95]{resizegather} % For resizing gather environment to fit massive equations
\setbeameroption{show notes} % Shows notes on the right of the main slides
% \setbeameroption{show notes on second screen}
\usefonttheme[onlymath]{serif} % For serif font in math mode
\usepackage[round]{natbib} % Needed for Author name citations
\usepackage{../../packages/shared}
\usepackage{../../packages/misc_commands}
\input{../../figures/causal_structures/formatting} % Causal structure formatting for tikz
\renewcommand{\term}[1]{\textcolor{Mahogany}{#1}}

@@

@@

\begin{frame}
    \frametitle{Incidence Matrix}
    \begin{itemize}
        \item Incidence matrix $M$ is a bit-wise matrix
        \item Row-indexed by marginal events $m \in \Events{\mscenario}$
        \item Column-indexed by joint events $j \in \Events{\jointvar}$
        \[ M_{m,j} = \begin{cases}
            1 & m \text{ compatible with } j \\
            0 & \text{otherwise}
        \end{cases} \]
    \end{itemize}
\end{frame}

\begin{frame}
    \frametitle{Application to Large Inflation}
    \begin{columns}
        \column{0.6\linewidth}
            \begin{itemize}
                \item Tackling Large inflation
                \item 12 pre-injectable sets (to follow)
            \end{itemize}
        \column{0.4\linewidth}
            \scalebox{0.5}{\input{../../figures/causal_structures/inflated_triangle_scenario}}
    \end{columns}
\end{frame}

@@

\begin{frame}
    \frametitle{Large Inflation Incidence}
    \begin{itemize}
        \item Joint variables are all of the observable nodes $\nodes_O' = \jointvar$
        \[ \jointvar = \bc{A_1, A_2, A_3, A_4,B_1, B_2, B_3, B_4,C_1, C_2, C_3, C_4} \]
        \item Marginal scenario is composed of pre-injectable sets $\mscenario = \PreInj[\graph]{\graph'}$
        \item Inequalities Violated by Fritz distribution are inherently $4$-outcome
        \item Incidence matrix $M$ is very large $\sim 2.25 \textsf{Gb}$
        \[ \# \text{Columns} = \prod_{v \in \jointvar} O_{v} = 4^{12} = 16,777,216 \]
        \[ \# \text{Rows} = \sum_{V \in \mscenario} \prod_{v \in V} O_{v} = 4\times 4^{6} + 8 \times 4^{3} = 16,896 \]
    \end{itemize}
\end{frame}

@@

\begin{frame}
    \frametitle{Parameterizing Unitary Group}
    \begin{itemize}
        \item Spengler, Huber and Heismayr \cite{Spengler_2010_Unitary} demonstrate a parameterization of $\s{U}\br{d}$ where the parameters are organized in a $d \times d$-matrix of real values $\la_{n,m}$
        \[ U = \bs{\prod_{m=1}^{d-1} \br{\prod_{n=m+1}^{d} R_{m,n} RP_{n,m}}} \tcdot \bs{\prod_{l=1}^{d} GP_{l}} \]
        \item Global Phase Terms: $GP_{l} = \exp\br{iP_l \lambda_{l,l}}$
        \item Relative Phase Terms: $RP_{m,n} = \exp\br{i P_n \lambda_{n,m}}$
        \item Rotation Terms: $R_{n,m} = \exp\br{i \si_{m,n} \lambda_{m,n}}$
        \item Projection Operators: $P_l = \ket{l}\bra{l}$
        \item Anti-symmetric $\si$-matrices: $\si_{m,n} = - i\ket{m}\bra{n} + i \ket{n}\bra{m}$
    \end{itemize}
\end{frame}

\begin{frame}
    \frametitle{Parameterizing Unitary Group Cont'd}
    \begin{itemize}
        \item Each parameter $\la_{n,m}$ has physical interpretation
        \item Degeneracies are easily eliminated such as global phase
        \[ \forall l = 1, \ldots, d : \la_{l,l} = 0 \implies GP_{l} = 1 \]
        \item Parameterize $U \in \s{U}\br{d}$ up to global phase denoted $\ti{U} \in \s{U}\br{d}$
        \item Computationally efficient
        \begin{alignat*}{2}
            GP_l &= \ident &&+ P_l \br{e^{i\la_{l,l}} - 1} \\
            RP_{m,n} &= \ident &&+ P_n \br{e^{i\la_{n,m}} - 1} \\
            R_{n,m} &= \ident &&+ \br{\ket{m}\bra{m} + \ket{n}\bra{n}} \br{\cos\lambda_{n,m} - 1} \\
            & &&+ \br{\ket{m}\bra{n} - \ket{n}\bra{m}} \sin\lambda_{n,m} \\
        \end{alignat*}
    \end{itemize}
\end{frame}

>>>>

% \usepackage{pgfpages} % For viewing slides on the right
\usepackage[warningthreshold=0.95]{resizegather} % For resizing gather environment to fit massive equations
\setbeameroption{show notes} % Shows notes on the right of the main slides
% \setbeameroption{show notes on second screen}
\usefonttheme[onlymath]{serif} % For serif font in math mode
\usepackage[round]{natbib} % Needed for Author name citations
\usepackage{../../packages/shared}
\usepackage{../../packages/misc_commands}
\input{../../figures/causal_structures/formatting} % Causal structure formatting for tikz
\renewcommand{\term}[1]{\textcolor{Mahogany}{#1}}
\renewcommand{\tcdot}{\cdot} % Tentative cdot, not sure if I want to keep dots

@@

\begin{frame}
    \frametitle{$d$-Separation Polynomial}

@@

\begin{frame}
    \frametitle{Distribution Vectors}
    \begin{itemize}
        \item \term{Joint Distribution Vector $\probvec^{\jointvar}$}
        \[ \forall j \in \Events{\jointvar} : \probvec^{\jointvar}_j = \prob[\jointvar][j] \]
        \item \term{Marginal Distribution Vector $\probvec^{\mscenario}$}
        \[ \forall m \in \Events{\mscenario} : \probvec^{\mscenario}_m = \prob[\Dom{m}][m] \]
    \end{itemize}
\end{frame}

\begin{frame}
    \frametitle{Incidence Matrix}
    \begin{itemize}
        \item Marginal Problem as linear program
        \begin{align*}
            \probvec^{\mscenario}_m &= {\sum}_{j} M_{m,j} \probvec^{\jointvar}_j \\
            \probvec^{\mscenario} &= M \tcdot \probvec^{\jointvar}
        \end{align*}
        \item \term{Incidence matrix} $M$ is a bit-wise matrix
        \item Row-indexed by marginal events $m \in \Events{\mscenario}$
        \item Column-indexed by joint events $j \in \Events{\jointvar}$
        \[ M_{m,j} = \begin{cases}
            1 & m = j|_{\Dom{m}} \\
            0 & \text{otherwise}
        \end{cases} \]
        \begin{gather*}
        \# \text{Columns} = \abs{\Events{\jointvar}} = \prod_{v \in \jointvar} \abs{O_{v}} \\
        \# \text{Rows} = \abs{\Events{\mscenario}} = \sum_{V \in \mscenario} \prod_{v \in V} \abs{O_{v}}
        \end{gather*}

    \end{itemize}
\end{frame}

\begin{frame}[shrink=10]
    \frametitle{Example}
    Let $\jointvar$ be $3$ binary variables $\jointvar = \bc{A,B,C}$ and $\mscenario$ be the marginal scenario $\mscenario = \bc{\bc{A,B}, \bc{B,C}, \bc{A,C}}$. The incidence matrix becomes:
    \[ M = \kbordermatrix{
        (A,B,C) \:\: = & (0,0,0) & (0,0,1) & (0,1,0) & (0,1,1) & (1,0,0) & (1,0,1) & (1,1,0) & (1,1,1) \\
        (A=0, B=0) & \kone & \kone & \kzer & \kzer & \kzer & \kzer & \kzer & \kzer \\
        (A=0, B=1) & \kzer & \kzer & \kone & \kone & \kzer & \kzer & \kzer & \kzer \\
        (A=1, B=0) & \kzer & \kzer & \kzer & \kzer & \kone & \kone & \kzer & \kzer \\
        (A=1, B=1) & \kzer & \kzer & \kzer & \kzer & \kzer & \kzer & \kone & \kone \\
        (B=0, C=0) & \kone & \kzer & \kzer & \kzer & \kone & \kzer & \kzer & \kzer \\
        (B=0, C=1) & \kzer & \kone & \kzer & \kzer & \kzer & \kone & \kzer & \kzer \\
        (B=1, C=0) & \kzer & \kzer & \kone & \kzer & \kzer & \kzer & \kone & \kzer \\
        (B=1, C=1) & \kzer & \kzer & \kzer & \kone & \kzer & \kzer & \kzer & \kone \\
        (A=0, C=0) & \kone & \kzer & \kone & \kzer & \kzer & \kzer & \kzer & \kzer \\
        (A=0, C=1) & \kzer & \kone & \kzer & \kone & \kzer & \kzer & \kzer & \kzer \\
        (A=1, C=0) & \kzer & \kzer & \kzer & \kzer & \kone & \kzer & \kone & \kzer \\
        (A=1, C=1) & \kzer & \kzer & \kzer & \kzer & \kzer & \kone & \kzer & \kone
    } \]
\end{frame}

\begin{frame}
    \frametitle{Application to Large Inflation}
    \begin{columns}
        \column{0.6\linewidth}
            \begin{itemize}
                \item Find compatibility inequalities for TS by finding non-contextuality inequalities for Large Inflation
                \item Marginal Scenario is the set of maximal pre-injectable sets
                \item 12 maximal pre-injectable sets (to follow)
            \end{itemize}
        \column{0.4\linewidth}
            \scalebox{0.5}{\input{../../figures/causal_structures/inflated_triangle_scenario}}
    \end{columns}
\end{frame}

@@

\begin{frame}
    \frametitle{Large Inflation Incidence}
    \begin{itemize}
        \item Joint variables are all of the observable nodes $\nodes_O' = \jointvar$
        \[ \jointvar = \bc{A_1, A_2, A_3, A_4,B_1, B_2, B_3, B_4,C_1, C_2, C_3, C_4} \]
        \item Marginal scenario is composed of pre-injectable sets $\mscenario = \PreInj[\graph]{\graph'}$
        \item Inequalities Violated by Fritz distribution are inherently $4$-outcome
        \item Incidence matrix $M$ is very large $\sim 2.25 \textsf{Gb}$
        \[ \# \text{Columns} = 4^{12} = 16,777,216 \]
        \[ \# \text{Rows} =  4\times 4^{6} + 8 \times 4^{3} = 16,896 \]
    \end{itemize}
\end{frame}

@@

\begin{frame}
    \frametitle{Parameterizing Unitary Group}
    \begin{itemize}
        \item Spengler, Huber and Heismayr \cite{Spengler_2010_Unitary} demonstrate a parameterization of $\s{U}\br{d}$ where the parameters are organized in a $d \times d$-matrix of real values $\la_{n,m}$
        \[ U = \bs{\prod_{m=1}^{d-1} \br{\prod_{n=m+1}^{d} R_{m,n} RP_{n,m}}} \tcdot \bs{\prod_{l=1}^{d} GP_{l}} \]
        \item Global Phase Terms: $GP_{l} = \exp\br{iP_l \lambda_{l,l}}$
        \item Relative Phase Terms: $RP_{n,m} = \exp\br{i P_n \lambda_{n,m}}$
        \item Rotation Terms: $R_{m,n} = \exp\br{i \si_{m,n} \lambda_{m,n}}$
        \item Projection Operators: $P_l = \ket{l}\bra{l}$
        \item Anti-symmetric $\si$-matrices: $\si_{m,n} = - i\ket{m}\bra{n} + i \ket{n}\bra{m}$
    \end{itemize}
\end{frame}

\begin{frame}
    \frametitle{Parameterizing Unitary Group Cont'd}
    \begin{itemize}
        \item Each parameter $\la_{n,m}$ has physical interpretation
        \item Degeneracies are easily eliminated such as global phase
        \[ \forall l = 1, \ldots, d : \la_{l,l} = 0 \implies GP_{l} = 1 \]
        \item Parameterize $U \in \s{U}\br{d}$ up to global phase denoted $\ti{U} \in \s{U}\br{d}$
        \item Computationally efficient
        \begin{alignat*}{2}
            GP_l &= \ident &&+ P_l \br{e^{i\la_{l,l}} - 1} \\
            RP_{n,m} &= \ident &&+ P_n \br{e^{i\la_{n,m}} - 1} \\
            R_{m,n} &= \ident &&+ \br{\ket{m}\bra{m} + \ket{n}\bra{n}} \br{\cos\lambda_{n,m} - 1} \\
            & &&+ \br{\ket{m}\bra{n} - \ket{n}\bra{m}} \sin\lambda_{n,m} \\
        \end{alignat*}
    \end{itemize}
\end{frame}
